TCP Congestion Algorithms In Datacenters \\ \vspace{0.3cm} {\large LS Project}

### 2017-12-14 paper/paper.tex
Minor change
paper: updated+improved experimental setup description, and other minor changes

One scenario will be used to compare the performance of the TCP congestion algorithms:

\begin{enumerate}
	\item The user owns multiple endpoints in the network. In this case a congestion algorithm is applied at both the client and server side. This scenario applies to multitentant datacenter situations, where a user owns a server in the datacenter network and a node outside of the datacenter network. In this test the clients are at the receiving side and the user server and random server at the sending side. The tests will be performed from a multitenant datacenter perspective.
\end{enumerate}

The created testing environment is shown in figure \ref{fig:setup1}. The clients are located on one physical device and are listening each on a different port, which is set by the controller described in section [TO DO: ENTER SECTION]. The servers are separate physical devices as one of the servers possible runs Windows when a TCP congestion control algorithm is only available for Windows. The software used for the servers and clients is Ubuntu 17.04. If Windows is used, the server will be running Windows 10.

@@

In scenario 1 the user server uses the same algorithm as the first client and the random server uses the same algorithm as the second client. Tests in this scenario cover behavioral differences between TCP congestion algorithms when a user owns a node in the network of a multitenant datacenter.

In scenario 1 the TCP congestion algorithm is tested for fairness in sharing bandwidth with other TCP connections using other TCP congestion algorithms. In order to test the fairness of congestion algorithms, the bandwidth allocation per second per algorithm is measured. It can be determined whether a specific TCP congestion algorithm in combination with other TCP congestion algorithms results in the allocation of more or less bandwidth to that specific TCP connection.

The tests conducted on the algorithms have two environment variables: packet loss and delay. Those environment variables are simulated using \texttt{netem}, which is a network emulation functionality in several Linux distributions \cite{linux-netem}. \texttt{netem} is applied to the incoming side of the interfaces of the clients to emulate the same packet loss and delay. Windows, which is tested as a server, does not support \texttt{netem}. Therefore the packet loss and delay are applied to the incoming interfaces of the clients. Packet loss and delay are chosen as the behavior of most TCP algorithms depends on those variables [SRC]. The values for the environment variables are chosen based on measurements performed by Verizon \cite{verizon-latency}.

@@

TO DO: Type some text here.

@@

\subsubsection{Theoretical Algorithm Characteristics}

@@

The CTCP algorithm is a combination of a loss-based and delay-based congestion protocol \cite{compound-tcp-congestion}. For the loss-based component, the variable conventional congestion window, $cwnd$, was introduced. For the delay-based component, the variable delay window, $dwnd$, was introduced. The TCP sending window is determined from: $win = min(cwnd + dwnd, awnd)$, where $awnd$ is the window advertised by the receiver. For $cwnd$ on the arrival of an ACK: $cwnd = cwnd + 1/win$. The $dwnd$ variable is initially set to zero and is only set when the congestion avoidance is active. $dwnd$ is calculated differently for several scenarios. If the network path is underutilized: $dwnd(t+1) = dwnd(t) + \alpha*dwnd(t)^{k} - 1$. When the network path is congested: $dwnd(t+1) = dwnd(t) - \eta*diff$. When packet loss occurs: $dwnd(t+1) = dwnd(t)(1-\beta) - cwnd/2$. $k$, $\alpha$, $\eta$ and $\beta$ are tunable parameters for optimization of the algorithm. $\eta$ defines the rapidness of reduction of the window when congestion is detected. $diff$ is calculated from the expected throughput and the actual throughput, which are again calculated from the window size and Round Trip Time (RTT).

@@

\subsubsection{Performance comparison}

- Includes results of tests.

>>>>

We will analyse one common scenario: multiple servers of different owners
inside a multitenant datacenter, sends data to clients outside the datacenter.
This could be users downloading software, the bursts while streaming a video
or movie, or large resources on a webpage.

We will create a bottleneck on the path to the user to see how different
congestion algorithms handle this. The connection will be full duplex and have
symmetrical speeds, to avoid any congestion in sending acknowledgements.
Because the server is sending the large volume of data and the client is
responding with the much smaller acknowledgements, the server will be the one
applying the congestion control. Therefore we vary the algorithm settings of
the server and not the client. The client will always be a GNU/Linux machine
using CUBIC, but we assume (due to the aforementioned reasons) that this does
not matter.

The bottleneck is created by having one receiving host, listening on different
ports, and two transmitting hosts. Each network interface is 1 gigabit per
second, thus the servers will have to limit their speed in order to avoid
flooding the client. This setup is shown in figure \ref{fig:setup1}.

We use three physical hosts as servers: two running Ubuntu 17.04 and one
running Microsoft Windows 10. In each experiment, only two hosts will be used.
The Windows system is used because there is no working CTCP
implementation---the default algorithm in Windows since Vista---available for a
modern Linux kernel.

@@

We will test the TCP congestion algorithm for fairness in sharing bandwidth
with other TCP connections using other congestion algorithms. In order to test
the fairness, the bandwidth usage per second per algorithm is measured. If two
different algorithms use similar amounts of bandwidth, they can be said to be
fair relative to each other.

The tests conducted with the algorithms have two environment variables: packet loss and delay. Those environment variables are simulated using \texttt{netem}, which is a network emulation functionality in several Linux distributions \cite{linux-netem}. \texttt{netem} is applied to the incoming side of the interfaces of the clients to emulate the same packet loss and delay. Windows, which is tested as a server, does not support \texttt{netem}. Therefore the packet loss and delay are applied to the incoming interfaces of the clients. Packet loss and delay are chosen as the behavior of most TCP algorithms depends on those variables [TODO: SRC]. The values for the environment variables are chosen based on measurements performed by Verizon \cite{verizon-latency}.

Each test is run at least twice. When results are conflicting or unexpected, we
ran a test more often and on different hardware.

@@

TODO: Type some text here.

@@

\subsection{Theoretical Algorithm Characteristics}

@@

The CTCP algorithm is a combination of a loss-based and delay-based congestion protocol \cite{compound-tcp-congestion}. For the loss-based component, the variable conventional congestion window, $cwnd$, was introduced. For the delay-based component, the variable delay window, $dwnd$, was introduced. The TCP sending window is determined from: $win = min(cwnd + dwnd, awnd)$, where $awnd$ is the window advertised by the receiver. For $cwnd$ on the arrival of an ACK: $cwnd = cwnd + 1/win$. The $dwnd$ variable is initially set to zero and is only set when the congestion avoidance is active. $dwnd$ is calculated differently for several scenarios. If the network path is underutilized: $dwnd(t+1) = dwnd(t) + \alpha*dwnd(t)^{k} - 1$. When the network path is congested: $dwnd(t+1) = dwnd(t) - \eta*diff$. When packet loss occurs: $dwnd(t+1) = dwnd(t)(1-\beta) - cwnd/2$. $k$, $\alpha$, $\eta$ and $\beta$ are tunable parameters for optimization of the algorithm. $\eta$ defines the rapidness of reduction of the window when congestion is detected. $\beta$ is the factor used for the multiplicative decrease after loss is detected. $diff$ is calculated from the expected throughput and the actual throughput, which are again calculated from the window size and Round Trip Time (RTT).

@@

\subsection{Experiments' Results}

TODO
